---
title: "Pharma Practice"
format: 
  html:
    theme: mint
editor: visual
editor_options: 
  chunk_output_type: console
---

This practice project is an exploration of pharmaceutical data analysis techniques using sample datasets obtained from the Harvard Dataverse at <https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/51B6NK>. The project focuses on applying Clinical Data Interchange Standards Consortium (CDISC) Study Data Tabulation Model (SDTM) standards to real-world clinical trial data, with the goal of creating Analysis Data Model (ADaM) datasets and generating statistical reports typical of those used in the pharmaceutical industry. This hands-on exercise serves as a foundation for developing practical skills in statistical programming within the pharma sector.

Load packages.

```{r}
#| message: false
library(tidyverse)
library(metacore)
library(metatools)
library(pharmaversesdtm)
library(admiral)
library(xportr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
```

Load data.

```{r}
#| message: false
meta <- 
  jsonlite::read_json("data/meta.json")

file_names <- list.files("data/", "*.csv", full.names = TRUE)

all_data <- 
  file_names %>% 
  map(read_csv) %>% 
  set_names(str_remove(basename(file_names), ".csv"))
```

Based on the JSON metadata provided, here's a short description of each file:

`ae.tab`: This contains Adverse Events data, which likely includes details about any undesirable experiences associated with the use of a medical product in a patient.

`dm.tab`: This file contains Subject Demographics data, detailing information about the trial subjects such as age, sex, race, and other demographic details.

`eg.tab`: This file includes Electrogardiogram data, which are records of the electrical activity of the heart over a period of time.

`lb.tab`: This contains Laboratory Findings data, providing information from laboratory tests conducted during the study.

`mh.tab`: This file holds Medical History data, containing historical health information about the subjects enrolled in the clinical study.

`pe.tab`: This contains Physical Examination data, documenting the findings from physical exams conducted on study subjects.

`vs.tab`: This file contains Vital Signs data, which includes measurements like blood pressure, heart rate, body temperature, and respiratory rate.

Each file is in a tab-separated values format and was created for the purpose of demonstrating the use of CDISC-SDTM formatted datasets.
